---
title: "Gapminder-Exploration"
output: github_document
---
```{r}
options(warn = -1) # supresses warnings
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(gapminder)
library(tidyverse)
```

## Data Reshaping Task

For the reshaping task, I chose activity number 3. In addition, I chose to compute the median life expectancy for each combination of continent and year. To reformat the data with columns for each of the continents, I used the the spread function.
```{r results = 'asis'}
df <- gapminder %>%
  group_by(continent, year) %>%
  summarize(medLifeExp = median(lifeExp)) %>% # get median life expectancy for each combination of year and continent
  spread(key = "continent", value = "medLifeExp") # creates a variable for each continent

knitr::kable(df) # print table

ggplot(df, aes(year, Asia)) + 
    geom_line() + # plots a line plot
    ggtitle("Asia Median Life Expectancy versus Year") + 
    xlab("Year") + 
    ylab("Median Life Expectancy")
```

A very simple task with this new data format is the ability to plot median life expectancy for some given continent versus year. Each continent is now treated as a quantitative variable, so we can use, for example, a line plot or scatterplot. As an example, I choose to plot the median life expectancy versus year for Asia, using a line splot.

## Join Task

For this task, I chose activity 1. I also experimented with the merge function. I chose to join the gapminder dataset with a new dataset containing 7 countries from the original gapminder dataset, with their capitols and currencies. I filtered by the year 2007 to ensure I don't have multiple copies of each country in my final dataset.
```{r results = 'asis'}
olddf <- gapminder %>%
  filter(year == 2007) # keep only data points from 2007 

newdf <- tribble(
  ~country, ~capitol, ~currency,
  'Germany', 'Berlin', 'Euro',
  'United Kingdom', 'London', 'Pound',
  'Italy', 'Rome', 'Euro',
  'France', ' Paris', 'Euro',
  'Japan', 'Tokyo', 'Yen',
  'China', 'Beijing', 'Yuan',
  'Korea, Rep.', 'Seoul', 'Won'
) # creates a new dataset with various countries and their capitols and currency

df <- inner_join(olddf, newdf, by = 'country') # joins both created tables by country
dfMerge <- merge(olddf, newdf, by.x = 'country') # joins tables again, except with different function

knitr::kable(df) # print table created with join
knitr::kable(dfMerge) # print table created with merge
```

I was able to get the exact same tables using the inner join and merge functions. Instead of the inner join function, I could have also used the right join function, and gotten the same result, assuming I treat the original gapminder dataset as the left dataset and the new dataset as the right dataset. This is because the new dataset I created only contained a subset of the countries in the gapminder dataset. Using the left join, however, would have resulted in a dataset with many rows containing n/a values, unless I treated the new dataset as the left dataset.
