---
title: "Gapminder-Exploration"
output: github_document
---
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r cars}
library(gapminder)
library(tidyverse)
```

## Data Reshaping Task

For the reshaping task, I chose activity number 3. In addition, I chose to compute the median life expectancy for each combination of continent and year. To reformat the data with columns for each of the continents, I used the the spread function.
```{r results = 'asis'}
df <- gapminder %>%
  group_by(continent, year) %>%
  summarize(medLifeExp = median(lifeExp)) %>%
  spread(key = "continent", value = "medLifeExp") 
knitr::kable(df, 'markdown') # print table
ggplot(df, aes(year, Asia)) + 
    geom_line() +
    ggtitle("Asia Median Life Expectancy versus Year") + 
    xlab("Year") + 
    ylab("Median Life Expectancy")
```

A very simple task with this new data format is the ability to plot median life expectancy for some given continent versus year. Each continent is now treated as a quantitative variable, so we can use, for example, a line plot or scatterplot. As an example, I choose to plot the median life expectancy versus year for Asia.

## Join Task

```{r}

```